---
title: "Fired cops"
output: html_notebook
---


```{r echo=FALSE, results='hide', message=FALSE}
library(RPostgreSQL)
library(knitr)
options(scipen=999,digits=2)
options(dplyr.summarise.inform = FALSE)
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, max.print = -1)
Sys.setenv(PGPASSWORD="xxxx")
con <- dbConnect(RPostgres::Postgres(),
                     dbname="xxxx",
                 port="xxxx",
                 host="xxxx",
                 user="xxxx"
                 )
library(dplyr,warn.conflicts = FALSE)
library(tidyr)
library(tidyverse) 
library(ggplot2)
library(lubridate)
library(readr)
library(readxl)
library(fuzzyjoin)

```

## Identifying and grouping lists of fired and rehired cops from 2000 on

```{sql connection=con, output.var="fired_rehired_cops_appt", echo=FALSE, cache = TRUE}
SELECT * FROM
(SELECT * FROM 
(SELECT  fe.ptbid,
        fe.first,
        fe.last,
        fe.employer,
        fe.separation,
        fe.appointed,
        fe.separated,
        LEAD(fe.employer, 1) OVER (
        PARTITION BY fe.ptbid 
        ORDER BY fe.appointed ASC) next_cop_shop,
        LEAD(fe.appointed, 1) OVER (
        PARTITION BY fe.ptbid 
        ORDER BY fe.appointed ASC) appointment_after_fire
FROM 
(SELECT f.ptbid, MAX(f.separated) AS max_date FROM
(SELECT plur.ptbid,plur.ct FROM 
(SELECT ptbid, COUNT(*) ct FROM form_es
GROUP BY ptbid) AS plur 
WHERE plur.ct > 1) AS mult
JOIN form_es f USING (ptbid)
WHERE separation = 'Terminated for Cause'
GROUP BY f.ptbid) as term
JOIN form_es fe ON fe.ptbid = term.ptbid) as term
WHERE separation = 'Terminated for Cause' AND next_cop_shop IS NOT NULL) as every
WHERE appointment_after_fire >= separated AND appointed >= '2000-01-01'
```

```{r}
### remove cases where fired cop joined the same department twice
fired_rehired_cops_totals_appt <- fired_rehired_cops_appt %>% 
  group_by(next_cop_shop) %>%
  distinct(next_cop_shop, ptbid) %>% 
  ungroup() %>% 
  group_by(next_cop_shop) %>%
  summarize(total_fired_cops_rehired=n())

```


```{sql connection=con, output.var="all_cops_2000", echo=FALSE, cache = TRUE}
--- All cops from 2000 on
SELECT employer, COUNT(DISTINCT ptbid) as total_officers FROM form_es
WHERE appointed >= '2000-01-01'
GROUP BY employer
ORDER BY COUNT(DISTINCT ptbid) DESC

```

```{r}
#join data
total_pct <- all_cops_2000 %>%
  left_join(fired_rehired_cops_totals_appt,by=c("employer"="next_cop_shop")) %>% 
  mutate(pct=(total_fired_cops_rehired/total_officers)*100) %>% 
  filter(total_officers >= 100) %>% 
  arrange(-pct)
```

```{r}
#write to csv
write_csv(fired_rehired_cops_appt, '~/Desktop/robbins_story/all_fired_rehired_cops_2000_on.csv')
write_csv(total_pct, '~/Desktop/robbins_story/total_fired_hired_pct.csv')
```

### Appendix

This analysis was was performed on the "Form E" database of police officers (as of 2023-12-12) that is kept by the Illinois Training and Standards Board, which I obtained through a FOIA request. This database contains the name, birth year and individual key for every police officer in Illinois as well the dates of employment and the reason for their exit from every police department in the state.